---
title: Custom Feature Generation
author: Rado
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r echo=FALSE, warning = FALSE, message = FALSE}
# Clean workspace and load libs:
rm(list = ls())
library(plyr)
library(dplyr)
library(smbinning)
library(woeBinning)
library(ggplot2)
library(doParallel)
library(InformationValue)
source('_utils.R')

sdata <- read.csv('./sanitized_data/s_application_train.csv', header = T)
sdata <- castVariables(sdata)
```
# TO DOs:


# Intro

## Goal
Generation of features from loan application-level attributes

## Inputs
    - application_train.csv 
    - application_test.csv
    - ...

## Outputs
    - predictor_base_train.csv
    - predictor_base_test.csv

## Settings
```{r echo=TRUE, warning = FALSE, message = FALSE}
# Settings
pVal        <- 0.05 #pval for sm- and woe-binning
IVtresh     <- 0.05
minPerClass <- 0.01
```

# Overview of VARIABLE IMPORTANCE based on WOE and IV

```{r echo=FALSE, warning = FALSE, message = FALSE}
# --------------------------------------------------------------------------------------------------
#                           --- Quantitative variables ---
# --------------------------------------------------------------------------------------------------
sdata_q <- dplyr::select(sdata, starts_with("Q_")) %>%
  dplyr::mutate(T_TARGET = sdata$T_TARGET)

qnames            <- colnames(sdata_q)[colnames(sdata_q) != "T_TARGET"]
results_smBinning_q <- list()
results_WoeBinning_q <- list()

# RADO: The idea here is to run various 'optimal binning' algos, save results,
# then select only those variables and "cuts" from a given binning algo that 
# give IV larger than IVtresh.
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
nParts <- 10
log_con1 <- file("SmBinningLog.log", open = "a")
log_con2 <- file("WoeBinningLog.log", open = "a")
for(j in 1:nParts){ #this outer loop is there due to heap insuficiency
  print(paste0('Start: ', Sys.time()))
  minn <- (j-1)*ceiling(ncol(sdata_q)/nParts) + 1
  maxx <- min((j)*ceiling(ncol(sdata_q)/nParts),length(qnames))
  print(paste0('=== Chunk >> ', minn, ':', maxx, ' from ', length(qnames),'. ==='))
  
  aux_smBinning <- foreach(i = minn:maxx, .packages = c('smbinning')) %dopar% doSmBinning('q',qnames[[i]], sdata_q, pVal, IVtresh, './SmBinningLog.log')
  aux_woeBinning <- foreach(i = minn:maxx, .packages = c('woeBinning')) %dopar% doWoeBinning(qnames[[i]], sdata_q, pVal, IVtresh, minPerClass,'./WoeBinningLog.log')
  
  if(j == 1){
    results_smBinning_q   <- aux_smBinning
    results_WoeBinning_q  <- aux_woeBinning
  } else{
    results_smBinning_q   <- c(results_smBinning_q, aux_smBinning)
    results_WoeBinning_q  <- c(results_WoeBinning_q, aux_woeBinning)
  }
  results_smBinning_q <- plyr::compact(results_smBinning_q)
  results_WoeBinning_q <- plyr::compact(results_WoeBinning_q)
  
  rm(aux_smBinning, aux_woeBinning)
  print(paste0('End: ', Sys.time()))
}
stopCluster(cl)
close(log_con1)
close(log_con2)

# save the interesting vars from 'smBinning' to IV container
ivContainer       <- list()
for(res in results_smBinning_q){
    ivContainer[[colnames(sdata_q)[res$col_id]]][['sm']]$IV             <- res$iv
    ivContainer[[colnames(sdata_q)[res$col_id]]][['sm']]$cuts           <- res$cuts
    ivContainer[[colnames(sdata_q)[res$col_id]]][['sm']]$binningMethod  <- 'smBinning'
}

# save the interesting vars from 'WoeBinning' to IV container
for(res in results_WoeBinning_q){
    ivContainer[[res[[1]]]][['woe']]$IV             <- res[[2]]$iv.total.final[1] 
    ivContainer[[res[[1]]]][['woe']]$cuts           <- res[[2]]$cutpoints.final[-1]
    ivContainer[[res[[1]]]][['woe']]$binningMethod  <- 'woeBinning'
}

# --------------------------------------------------------------------------------------------------
#                           --- Categorical variables ---
# --------------------------------------------------------------------------------------------------
sdata_c <- dplyr::select(sdata, starts_with("CN_"), starts_with("CO_")) %>%
  dplyr::mutate(T_TARGET = sdata$T_TARGET)

cnames                <- colnames(sdata_c)[colnames(sdata_c) != "T_TARGET"]
results_smBinning_c   <- list()
results_WoeBinning_c  <- list()

# RADO: The idea here is to run various 'optimal binning' algos, save results,
# then select only those variables and "cuts" from a given binning algo that 
# give IV larger than IVtresh.
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
nParts <- 6
log_con1 <- file("SmBinningLog.log", open = "a")
log_con2 <- file("WoeBinningLog.log", open = "a")
for(j in 1:nParts){ #this outer loop is there due to heap insuficiency
  print(paste0('Start: ', Sys.time()))
  minn <- (j-1)*ceiling(ncol(sdata_c)/nParts) + 1
  maxx <- min((j)*ceiling(ncol(sdata_c)/nParts),length(cnames))
  print(paste0('=== Chunk >> ', minn, ':', maxx, ' from ', length(cnames),'. ==='))
  
  aux_smBinning <- foreach(i = minn:maxx, .packages = c('smbinning')) %dopar% doSmBinning('c',cnames[[i]], sdata_c, pVal, IVtresh, './SmBinningLog.log')
  aux_woeBinning <- foreach(i = minn:maxx, .packages = c('woeBinning')) %dopar% doWoeBinning(cnames[[i]], sdata_c, pVal, IVtresh, minPerClass,'./WoeBinningLog.log')
  
  if(j == 1){
    results_smBinning_c   <- aux_smBinning
    results_WoeBinning_c  <- aux_woeBinning
  } else{
    results_smBinning_c   <- c(results_smBinning_c, aux_smBinning)
    results_WoeBinning_c  <- c(results_WoeBinning_c, aux_woeBinning)
  }
  results_smBinning_c <- plyr::compact(results_smBinning_c)
  results_WoeBinning_c <- plyr::compact(results_WoeBinning_c)
  
  rm(aux_smBinning, aux_woeBinning)
  print(paste0('End: ', Sys.time()))
}
stopCluster(cl)
close(log_con1)
close(log_con2)

# save the interesting vars from 'smBinning' to IV container
for(res in results_smBinning_c){
    ivContainer[[colnames(sdata_c)[res$col_id]]][['sm']]$IV             <- res$iv
    ivContainer[[colnames(sdata_c)[res$col_id]]][['sm']]$cuts           <- res$cuts
    ivContainer[[colnames(sdata_c)[res$col_id]]][['sm']]$binningMethod  <- 'smBinning'
}

# save the interesting vars from 'WoeBinning' to IV container
for(res in results_WoeBinning_c){
    ivContainer[[res[[1]]]][['woe']]$IV             <- res[[2]]$iv.total.final[1] 
    ivContainer[[res[[1]]]][['woe']]$cuts           <- res[[2]]$cutpoints.final[-1]
    ivContainer[[res[[1]]]][['woe']]$binningMethod  <- 'woeBinning'
}

# --------------------------------------------------------------------------------------------------
#                           --- Indicator variables ---
# --------------------------------------------------------------------------------------------------
sdata_i <- dplyr::select(sdata, starts_with("I_")) %>%
  dplyr::mutate(T_TARGET = sdata$T_TARGET)

inames                <- colnames(sdata_i)[colnames(sdata_i) != "T_TARGET"]
for(var in inames){
  print(paste0("Calculating Information Value for: ", var))
  IV <- InformationValue::IV(X=sdata_i[,var], Y=sdata_i[,'T_TARGET'], valueOfGood = 1)
  print(paste0('IV: ', as.character(IV[[1]])))
  if(IV[[1]] > IVtresh){
    ivContainer[[var]][['noBinning']]$IV <- IV[[1]]
    ivContainer[[var]][['noBinning']]$binningMethod <- 'no binning'
  }
}

# --------------------------------------------------------------------------------------------------
#                           --- Plot ---
# --------------------------------------------------------------------------------------------------
counter     <- 0
dataForPlot <- data.frame()
for(el in ivContainer){
  counter <- counter +1
  for(method in names(el)){
    dataForPlot <- rbind(dataForPlot, data.frame(variable = paste0(names(ivContainer)[counter], ' | ', method)
                                , IV = el[[method]]$IV))
  }
}

ggplot(data = dataForPlot, aes(x = reorder(variable, IV), y = IV)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```

## Detailed Results for those with IV > IVtresh
### SM Binning
```{r echo=FALSE, warning = FALSE, message = FALSE}
smNames <- names(ivContainer)[grepl('_sm',names(ivContainer))]
for(i in 1:length(results_smBinning_q)){
  par(mfrow=c(2,2))
  varName <- smNames[i]
  boxplot(sdata_q[,gsub("_sm",'',varName)]~sdata_q$T_TARGET,horizontal=T, frame=F, col="lightgray",main="Distribution")
  smbinning.plot(results_smBinning_q[[i]],option="dist")
  smbinning.plot(results_smBinning_q[[i]],option="badrate")
  smbinning.plot(results_smBinning_q[[i]],option="WoE")
  mtext(varName, side = 3, line = -13, outer = TRUE)
}
```
### WOE Binning
TODO: Dunno how to plot it into the document.
```{r echo=FALSE, warning = FALSE, message = FALSE}
woeNames <- names(ivContainer)[grepl('_woe',names(ivContainer))]
for(i in 1:length(results_WoeBinning)){
  woe.binning.plot(results_WoeBinning[[i]])
}
```


# Variable importance based on gradient boosting
TBD


# (bucketize and) Save
```{r echo=FALSE, warning = FALSE, message = FALSE}
## Final dataset
# 1) take target and id
output <- sdata[,c("T_TARGET", "ID_SK_ID_CURR")]

# 2) take original variables from the top IV-selected
interesting_ones <- names(ivContainer)
output <- cbind(output, sdata[,interesting_ones])

# 3.i) add SM binning of these
qnames_sm <- sapply(results_smBinning_q, function(y) y$x)
cnames_sm <- sapply(results_smBinning_c, function(y) y$x)

for(namee in qnames_sm){
  newName           <- paste0('B_',namee,'_sm')
  output[,newName]  <- paste0("Bin_", findInterval(sdata[,namee], results_smBinning_q[[which(qnames_sm == namee)]]$cuts))  
}

#TODO: Generating optimal bins of categorical variables from SMbinning output, to be figured out...

# 3.i) add WoE binning of these
qnames_woe <- sapply(results_WoeBinning_q, function(y) y[[1]])
cnames_woe <- sapply(results_WoeBinning_c, function(y) y[[1]])

for(namee in qnames_woe){
  newName           <- paste0('B_',namee,'_woe')
  # TODO: figure out how the edge points of the bins can be extracted from the WOE binning...
}

#TODO: Generating optimal bins of categorical variables from SMbinning output, to be figured out...

# save output
write.csv(output, file='./predictor_base/predictor_base_train.csv')
```
Output saved to '03_data_preparation/predictor_base'.